Collocated Deduplication with Fault Isolation for Virtual Machine Snapshot Backup

### 2013-05-30 snapshot_system/arch.tex
add deletion

For the Zipf-like distribution, an approximation to the sum of the first $n$ 
elements of the distribution can be derived as follows:
\begin{equation}
\sum_{i=1}^{n}\frac{1}{i^\alpha}\approx \int_{1}^{n}\frac{1}{x^\alpha}\mathrm{d}x=\frac{x^{1-\alpha}}{1-\alpha}=\frac{n^{1-\alpha}}{1-\alpha}\;  for\;  \alpha<1
\end{equation}
So the cumulative distribution function for a CDS holding top $S_c$ fingerprints
of global index with size $S_g$ is:
\begin{equation}
  E = (S_c / S_g)^{1-\alpha} \;  for\;  \alpha<1
\end{equation}
%[Describe the dedup efficiency model in detail]
Let $N$ be the number of nodes in the cluster, $m$ be the memory on each node that are used by CDS, $D$ be the amount of data on each node, and $B$ be the average block size. Then $S_c$ and $S_g$ can be expressed as:
\begin{equation}
S_c = N*m/F, S_g = N*D/B
\end{equation}
By replacing $S_c$ and $S_g$ in the first formula, the deduplication efficiency becomes:
\begin{equation}
  E = (\frac{m*B}{F*D})^{1-\alpha}
\end{equation}
Since $B$, $D$ and $F$ are pre-configured constants, the deduplication efficiency of CDS is only controlled by the its memory usage.

@@

%uses the logical a bloom filter to check if a data are still referenced by living snapshots. 
{\bf Discussion}
For one VM's snapshots, let $D_{leak}$ be the amount of storage leakage, 
$D_{del}$ be the average amount of data that should be physically deleted in one snapshot deletion,
$P_{1}$ be the false-positive rate of merged bloom filter type-1,
$N_1$ be the number of runs of approximate deletion phase-1, then we have:
\begin{equation}
D_{leak} = N_1 * P_b * D_{del}
\end{equation}

We let approxmite deletion phase-2 be triggered  
when $D_{leak}/D_{del}$ is accmulated to exceed certain threshold $t$:
\begin{equation}
\frac{D_{leak}}{D_{del}} = N_1 * P_1 > t \Rightarrow N_{1} > \frac{t}{P_1}
\end{equation}
For example, letting $P_1 = 0.01$ and $t=1.0$, 
we would expect deletion phase-2 be triggered once for 
every $T/P_{bl} = 100$ runs of deletion phase-1.
On a node that hosts 20 VMs and each VM deletes one snapshot per day, there would be 
only 1 deletion phase-2 scheduled for every 5 days, which is sufficiently small to prevent the 
heavy I/O workload of deletion phase-2 from disturbing normal VM opeations.

>>>>

For the Zipf-like distribution, an approximation to the sum of the first $n$ 
elements of the distribution can be derived as follows:
\begin{equation}
\sum_{i=1}^{n}\frac{1}{i^\alpha}\approx \int_{1}^{n}\frac{1}{x^\alpha}\mathrm{d}x=\frac{x^{1-\alpha}}{1-\alpha}=\frac{n^{1-\alpha}}{1-\alpha}\;  for\;  \alpha<1
\end{equation}
So the cumulative distribution function for a CDS holding top $S_c$ fingerprints
of global index with size $S_g$ is:
\begin{equation}
  E = (S_c / S_g)^{1-\alpha} \;  for\;  \alpha<1
\end{equation}
%[Describe the dedup efficiency model in detail]
Let $N$ be the number of nodes in the cluster, $m$ be the memory on each node that are used by CDS, $D$ be the amount of data on each node, and $B$ be the average block size. Then $S_c$ and $S_g$ can be expressed as:
\begin{equation}
S_c = N*m/F, \; S_g = N*D/B
\end{equation}
By replacing $S_c$ and $S_g$ in the first formula, the deduplication efficiency becomes:
\begin{equation}
  E = (\frac{m*B}{F*D})^{1-\alpha}
\end{equation}
Since $B$, $D$ and $F$ are pre-configured constants, the deduplication efficiency of CDS is only controlled by the its memory usage.

@@

%uses the logical a bloom filter to check if a data are still referenced by living snapshots. 
{\bf Discussion}
For one VM's snapshots, let $L_{temp}$ be the amount of temporary storage leakage, 
$D_{del}$ be the average amount of data that should be physically deleted in one snapshot deletion,
$P_{1}$ be the false-positive rate of merged bloom filter type-1,
$N_1$ be the number of runs of deletion phase-1 during the gap of two deletion phase-2 operations, then we have:
\begin{equation}
L_{temp} = N_1 * P_b * D_{del}
\end{equation}

If we let approxmite deletion phase-2 be triggered  
when $L_{temp}/D_{del}$ is accmulated to exceed certain threshold $t$, then:
\begin{equation}
\frac{L_{temp}}{D_{del}} = N_1 * P_1 > t \Rightarrow N_{1} > \frac{t}{P_1}
\end{equation}

Let $P_{2}$ be the false-positive rate of merged snapshot summaries type-2,
$L_{perm}$ be the permanent storage leakage resulting from the approxmite deletion phase-2,
it can be calculated as follows: 
\begin{equation}
L_{perm} = 
\end{equation}

For example, letting $P_1 = 0.01$ and $t=1.0$, 
we would expect deletion phase-2 be triggered once for 
every $T/P_{bl} = 100$ runs of deletion phase-1.
On a node that hosts 20 VMs and each VM deletes one snapshot per day, there would be 
only 1 deletion phase-2 scheduled for every 5 days, which is sufficiently small to prevent the 
heavy I/O workload of deletion phase-2 from disturbing normal VM opeations.

### 2013-05-30 snapshot_system/impl.tex
add deletion

\section{Implementation}
\subsection{Snapshot Backup}
\subsection{Snapshot Read}
\subsection{Snapshot Deletion}
In a mature VM cluster, snapshot deletions are as frequent as snapshot creations.
Our system adopts lazy delete strategy so that all snapshot deletions are scheduled
in the backup time window at midnight. 
Therefore, snapshot deletions must be fast enough to fit in time window and
efficient enough to satisfy our resource constraints.
However, there is no simple solution can achieve these goals with high reliability.
In Aliyun's cloud, we use a {\em fuzzy deletion} method to trade deletion accuracy for
speed and resource usage. This method tolerates a tiny percentage of storage leakage
to make the deletion operation faster and efficient by an order of magnitude,
yet we still adopt an slow but accurate deletion method to fix such leakage in the long-term.
Our hybrid deletion strategy, using fuzzy deletion regularly and accurate deletion periodically,
accomplishs our speed, resource usage and relibility goals very well.

{\bf Approximate Deletion}
We design approximate deletion to fast identify unused data in the append store.
Instead of scanning the whold append store, it uses a bloom filter to check 
if a data are still referenced by living snapshots. However, there is a small false-positive probility which
would indetify unused data as in use, i.e., storage leakage.
The following steps would take place during an approximate deletion:

@@

The design goal of approximate deletion is to reduce the frequency of running accurate deletion.
For one VM's snapshots, let $D_{leakage}$ be the amount of storage leakage, $D_{del}$ be the average amount of data
to be deleted in one snapshot deletion, 
then we have:
\begin{equation}
D_{leakage} = N_{approximate} * P_{bl} * D_{del}
\end{equation}
where $N_{approximate}$ is the number of runs of approximate deletion. 
An accurate deletion is triggered to fix the storage leakage
when $D_{leakage}/D_{del}$ is accmulated to exceed certain threshold $T$:
\begin{equation}
D_{leakage} / D_{del} = N_{approximate} * P_{bl} > T \Rightarrow N_{approximate} > T/P_{bl}
\end{equation}
Therefore, when $P_{bl} = 0.01$ and $T=1$, 
there would be a run of accurate deletion for every $T/P_{bl} = 100$ runs of approximate deletion.
On a machine that hosts 20 VMs and each VM deletes snapshot daily, there would be less than
one accurate deletion scheduled per day.

{\bf Append Store Cleanup}
Our accurate deletion uses mark-and-sweep to find all the unused chunks in a VM's append store.
The following steps would take place during an accurate deletion:

\begin{enumerate}
\item Scan all the containers. For each container, extract full list of existing data references and allocate a bitmap. 
(Notice this list is self sorted.)
\item Scan all the snapshot recipes and segment recipes. For every data reference pointing to append store,
find it in the above lists and mark the corresponding position in bitmaps.
\item Scan the bitmaps, for every bit that are not marked, tell append store the corresponding data reference can be deleted.
\end{enumerate}

Although accurate deletion guarantees reliability and correctness, its speed and resource usage are big problems to our VM cloud.
Let $T_{scan\_AS}$ be the time to scan the VM's append store, $T_{scan\_recipes}$ be the time to scan one snapshot recipe
and its segment recipes, $T_{mark}$ be the time of finding a data reference in the list and mark bitmap, 
$N_{SS}$ be the number of snapshots and $N_c$ be the number of chunks in one snapshot.
The overall time of running an accurate deletion for one VM is:
\begin{equation}
T = T_{scan\_AS} + N_{SS} * T_{scan\_recipes} + N_{SS} * N_c * T_{mark}
\end{equation}

For a virtual disk that has 10 snapshots, 50 GB of data in the append store, $T_{scan\_AS}$ would have cost half an hour already.
In addition, scanning recipes of 10 snapshots cost another 5 minutes.
Furthermore, maintaining the full lists of data references and bitmaps costs about 100 MB of memory. Having tens of VMs
co-lived in one physical machine and deleting snapshot on a daily basis, 
it's infeasible to run accurate deletion as frequent as snapshot deletions.

>>>>

\section{Implementation}
\subsection{Snapshot Backup}
\subsection{Snapshot Read}
\subsection{Snapshot Deletion}
The following steps would take place during an approximate deletion:

@@
